Basic Parameter estimation of Binary Neutron Star Systems by the Advanced LIGO/Virgo Network

### 2013-09-23 doc/paper.tex
final version that was posted on Astro-Ph

Finally, note that the PSD that defines the inner
product \eqref{innerProduct} is simply the time-averaged sensitivity
of a given detector to a specific frequency.  Ignoring non-Gaussian
glitches, any stretch of data should contain a specific noise
realization drawn from a Gaussian colored by the PSD.  Without a
realistic model including the actual instrument sensitivity, non-stationary
 noise, and uncertainty in the PSD we cannot accurately
simulate gravitational-wave data to the point of making precise predictions about
parameter estimation capabilities.  Furthermore, the posterior
distribution from a future detection will be influenced by the
realization of the noise at the time of the detection.  Because of
this, we elect to focus only on the statistical errors due to the
flexibility of the waveforms and the expected sensitivity of the
advanced era detectors and suppress any effects of detector noise by
picking $n(t) \equiv 0$.  It can be shown \cite{Nissanke2010} that this ``zero-noise'' injection
represents the geometric mean of the same experiment over many different
noise realizations drawn from the same PSD.

Consequently, what we recover with the MCMC could be interpreted as
the posterior distribution function for each source averaged over all
possible (Gaussian) noise realizations, assuming we have the correct
PSD and that the signal is sufficiently powerful for the linearized-signal
approximation to be valid.  By selecting a source population with an
 artificial distribution in SNR, we ensure that the
     ``zero-noise'' as average uncertainty approximation is valid.  To check that our
      reported statistical errors are not in gross
disagreement with those obtained in the presence of noise, we selected
a particular simulated signal to perform parameter estimation in
several noise realizations.  The comparison of the noise-dependent and
noise-free results can be found in Section \ref{noiseSection}.

@@

The difference in uncertainties between the two studies is illustrative.  \cite{Nissanke2013}
considered a realistic distribution of sources motivated by a local galaxy catalogue for
 sources where $D\leq 200\text{Mpc}$, and homogeneously distributed in co-moving
 volume for sources where $D > 200\text{Mpc}$.  These were then limited to those signals
 where $\text{SNR}_{network} \geq 8.5$ (the threshold for detection candidates in the
 current LIGO/Virgo matched-filtering framework) for the MCMC parameter estimation.
   The current study considers a source
 population isotropically distributed on the sky, but at distances such that each signal
 represents an optimal case for parameter estimation.   Fig. \ref{compSky} compares
 the sky-area uncertainties that will be available to those considering ``good 
 gravitational-wave events'' (e.g. precision source localization), and the 
 uncertainties that one can gather by considering all gravitational-wave candidates with 
 $\text{SNR}_{network}\geq 8.5$ (e.g. electromagnetic follow-up).  As expected, both source
 populations show a dramatic decrease in parameter estimation uncertainties with the
 addition of the LIGO India site to the advanced detector network.  This holds true even when
 considering the increased number of nominal sources available to the HLVI network 
 configuration in the Nissanke population, where the expansion of the network detection 
 horizon due to LIGO India increases the number of sources to 130, versus 90 in 
 the HLV configuration.  This trend is similar to that observed in \cite{Veitch2012}, contrasting
 the originally proposed four-detector configuration (with two co-located Hanford detectors) to 
 the HLVI configuration.

@@

The authors would like to thank Samaya Nissanke, John Veitch, and Neil Cornish for useful discussions.  We would also like to thank Samaya Nissanke for the 95\% credible regions
quoted in Fig. \ref{compSky}. CR and BF were supported by NSF GRFP Fellowships, award DGE-0824162.

>>>>

Finally, note that the PSD that defines the inner
product \eqref{innerProduct} is simply the time-averaged sensitivity
of a given detector to a specific frequency.  Ignoring non-Gaussian
glitches, any stretch of data should contain a specific noise
realization drawn from a Gaussian colored by the PSD.  Without a
realistic model including the actual instrument sensitivity, non-stationary
 noise, and uncertainty in the PSD we cannot accurately
simulate gravitational-wave data to the point of making precise predictions about
parameter estimation capabilities.  Furthermore, the posterior
distribution from a future detection will be influenced by the
realization of the noise at the time of the detection.  Because of
this, we elect to focus only on the statistical errors due to the
flexibility of the waveforms and the expected sensitivity of the
advanced era detectors and suppress any effects of detector noise by
picking $n(t) \equiv 0$.

Consequently, what we recover with the MCMC could be interpreted as
the posterior distribution function for each source averaged over all
possible (Gaussian) noise realizations, assuming we have the correct
PSD and that the signal is sufficiently powerful for the linearized-signal
approximation to be valid.  By selecting a source population with an
 artificial distribution in SNR, we ensure that the
     ``zero-noise'' as average uncertainty approximation is valid.
       This approach has been employed in previous studies to similar effect
        (e.g. \cite{Nissanke2010} and \cite{Sampson2013}).
 To check that our reported statistical errors are not in gross
disagreement with those obtained in the presence of noise, we selected
a particular simulated signal to perform parameter estimation upon in
several noise realizations.  The comparison of the noise-dependent and
noise-free results can be found in Section \ref{noiseSection}.

@@

The difference in uncertainties between the two studies is illustrative.  \cite{Nissanke2013}
considered a realistic distribution of sources motivated by a local galaxy catalogue for
 sources where $D\leq 200\text{Mpc}$, and homogeneously distributed in co-moving
 volume for sources where $D > 200\text{Mpc}$.  These were then limited to those signals
 where $\text{SNR}_{network} \geq 8.5$ (the threshold for detection candidates in the
 current LIGO/Virgo matched-filtering framework) for the MCMC parameter estimation.
   The current study considers a source
 population isotropically distributed on the sky, but at distances such that each signal
 represents an optimal case for parameter estimation.   Fig.\ \ref{compSky} compares
 the sky-area uncertainties that will be available to those considering ``good 
 gravitational-wave events'' (e.g. precision source localization), and the 
 uncertainties that one can gather by considering all gravitational-wave candidates with 
 $\text{SNR}_{network}\geq 8.5$ (e.g. electromagnetic follow-up).  As expected, both source
 populations show a dramatic decrease in parameter estimation uncertainties with the
 addition of the LIGO India site to the advanced detector network.  This holds true even when
 considering the increased number of nominal sources available to the HLVI network 
 configuration in the Nissanke population, where the expansion of the network detection 
 horizon due to LIGO India increases the number of sources to 130, versus 90 in 
 the HLV configuration.  This trend is similar to that observed in \cite{Veitch2012}, contrasting
 the originally proposed four-detector configuration (with two co-located Hanford detectors) to 
 the HLVI configuration.

@@

The authors would like to thank Samaya Nissanke, John Veitch, and Neil Cornish for useful discussions.  We would also like to thank Samaya Nissanke for the 95\% credible regions
quoted in Fig. \ref{compSky}. CR and BF were supported by NSF GRFP Fellowships, award DGE-0824162.  All authors were also partially or fully supported by NSF Gravitational Physics grant PHY-0969820, PI: VK. Lastly, VK is grateful to the hospitality of the Aspen Center for Physics while she worked on this study.
